WORCS: A Workflow for Open Reproducible Code in Science

### 2020-10-31 paper/Manuscript_DS.Rmd
Rewrite reproducibility section

Many solutions exist to ensure computational reproducibility, which differ in
user-friendliness and effectiveness. 
These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs.
If strict reproducibility is required, users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
virtual computer, which can be stored like a time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the `R` environment (e.g.,
preprocessing tools for neuroimaging data).
The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
describes how users may instantiate a Docker build running `worcs` and its dependencies.
For more information on this topic,
@peikertReproducibleDataAnalysis2019 describe 
a Docker-based workflow for computational reproducibility that is
conceptually and practically compatible with WORCS.
One remaining challenge of Docker is that it is somewhat difficult to set up for novice users.
The `worcs` package meets most reproducibility requirements while using only tools
native to RStudio, and is therefore easier to adopt for `R` users.
However, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Code Ocean'. Although Code Ocean enjoys some popularity among 
scientists, it is not clear whether this platform
is GDPR compliant - as of this writing, its [terms of use](https://codeocean.com/terms-of-use) do not address the GDPR, nor is it clear where
data are stored.
The necessity to upload human participant data is a liability
when using any
cloud-computing solution for data analysis.

For analyses conducted entirely in `R`, the recently released package `renv`
strikes a great balance between user-friendliness  and effectiveness. Developed by
the team behind RStudio, this package simply maintains a text-based,
human-readable log of all packages used, their version numbers, and where they
were installed from (e.g., CRAN, Bioconductor, GitHub). As long as these
repositories maintain their archives, this solution is reliable and very
lightweight.
The text-based log file can be version controlled with Git. `renv`
installs the logged packages in a cache that is shared between all
projects on the computer. Thus, if one of your projects uses a package that is
already in the cache, it does not
need to be installed again. This prevents long installation
times, large space requirements, and frustration having to re-install software
that is already installed on the system. When someone else loads your 
project, `renv` will install all of the required packages onto a cache on their
computer. As opposed to the alternatives `renv` is tightly integrated with the
native R package management system and thus does not require the users to change their workflow of installing/removing and using packages.
This is not a fail-safe method to ensure strict computational reproducibility, but it is more user-friendly
and lightweight than any alternative out there, requiring no technical skill to
setup. When choosing an appropriate solution for dependency management, it is
important to consider that all solutions have a limited shelf life
[@brownHowLearnedStop2017].
Ultimately, the best approach might be to use a "good-enough" solution, and
acknowledge that all code requires some maintenance if you want to
reproduce it in the future. For projects in `R`, `renv` meets these requirements.

>>>>

Many solutions exist to ensure computational reproducibility,
which differ in user-friendliness and effectiveness.
These solutions typically work by 
enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs.
When choosing an appropriate solution for dependency management, there is a tradeoff between ease-of-use on the one hand, and robustness on the other.
In making this tradeoff, it is important to consider that all solutions have a limited shelf life
[@brownHowLearnedStop2017].
Therefore, the best approach might be to use a "good-enough" solution, and
acknowledge that all code requires some maintenance if you want to
reproduce it in the future.

For analyses conducted entirely in `R`, the recently released package `renv` [@usheyRenvProjectEnvironments2020]
strikes a great balance between user-friendliness and robustness.
Developed by the team behind RStudio, 
this package simply maintains a text-based,
human-readable log of all packages used, their version numbers, and where they
were installed from (e.g., CRAN, Bioconductor, GitHub).
This lightweight approach safeguards reproducibility as long as these repositories maintain their archives.
The text-based log file can be version controlled with Git.
The `renv` function `snapshot()` is used to add packages to the log file.
This function identifies all packages loaded using `library()` calls in the scripts in the project directory.

All solutions to computational reproducibility require users to make their dependencies explicit, often by reinstalling all software for each project in a controlled environment.
This can lead to long installation times and large memory requirements.
Similarly, `renv` requires users to call `install.packages()` for each library used - but it does not reinstall packages for each new project.
Instead, packages reside in a cache that is shared between all projects on the computer.
If a project uses a package that is already in the cache, it is not reinstalled, but linked from the cache.
This improves the user experience by saving time and memory.
When someone else loads your project, `renv` will install all of the required packages.
As opposed to alternative solutions, `renv` is tightly integrated with the
native R package management system.
Therefore, it does not require the users to change their workflow of installing/removing and using packages, and requires no technical skill to setup.
It is not a fail-safe method to ensure strict computational reproducibility,
but it meets most reproducibility requirements using only tools
native to RStudio, and is therefore easier to adopt for `R` users.

If stricter reproducibility is required,
users might choose to conduct all analyses in a "Docker container": An environment that behaves like a
virtual computer, which can be stored like a time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the `R` environment (e.g.,
preprocessing tools for neuroimaging data).
The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
describes how to instantiate a Docker build running `worcs` and its dependencies.
For more information on this topic,
@peikertReproducibleDataAnalysis2019 describe 
a Docker-based workflow for computational reproducibility that is
conceptually and practically compatible with WORCS.
One remaining challenge of Docker is that it is somewhat difficult to set up for novice users.
To overcome this challenge, an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Code Ocean'. Although Code Ocean enjoys some popularity among 
scientists, it is not clear whether this platform
is GDPR compliant - as of this writing, its [terms of use](https://codeocean.com/terms-of-use) do not address the GDPR, nor is it clear where
data are stored.
The necessity to upload human participant data is a liability
when using any
cloud-computing solution for data analysis.
